---
title: "Package Ecosystem"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package Ecosystem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## What is a package?

An R package is a collection of functions that are bundled together in a way that lets them be easily shared. Usually these functions are designed to work together to complete a specific task such as analysing a particular kind of data. You are probably familiar with many packages already, for example ggplot2 or dplyr.

Packages can take various forms during their life cycle. For more details have a look at the “What is a package?” section of Hadley Wickham’s “R packages” book (http://r-pkgs.had.co.nz/package.html#package)


## Why write a package?

Packages are the best way to distribute code and documentation, and as we are about to find out they are very simple to make. Even if you never intend to share your package it is useful to have a place to store your commonly used functions. You may have heard the advice:

> If you find yourself reusing code then you should wait anymore to turn it into a function so that you don’t have to keep rewriting it (_along with other benefits..._). 

The same applies to functions. If you have some functions you reuse in different projects then it probably makes sense to put those in a package. It’s a bit more effort now but it will save you a lot of time in the long run.

Of course often you will want to share your package, either to let other people use your functions or just so people can see what you have done (for example when you have code and data for a publication). If you are thinking about making a software package for public use there are a few things you should consider first:

 *  Is your idea new or is there already a package out there that does something similar? If you have an idea, [raise an issue to start the discussion asap](https://github.com/unhcRverse/unhcRverse/issues)! 
 
 *  If there is does your package improve on it in some way? For example is it easier to use or does it have better performance? Maybe check the existing packages and see if you can improve first the existing solution before starting a new one from scratch. UNHCR package developers are open to collaboration and you may be able to achieve more by working together.


## How to develop packages?

> We **DO STRONGLY RECOMMEND** to use [fusen package](https://thinkr-open.github.io/fusen/index.html) to develop packages as it facilitates a lot peer review and enforces a series of good practices (example, unit testing..)

In addition, there are plenty of resources available to learn how to develop packages. See a list below: 

 *  Writing an R package from scratch (Hilary Parker) https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/
 *  Writing an R package from scratch (Updated) (Thomas Westlake) https://r-mageddon.netlify.com/post/writing-an-r-package-from-scratch/
 *  usethis workflow for package development (Emil Hvitfeldt) https://www.hvitfeldt.me/blog/usethis-workflow-for-package-development/
 *  R package primer (Karl Broman) https://kbroman.org/pkg_primer/
 *  R Package Development Pictorial (Matthew J Denny) http://www.mjdenny.com/R_Package_Pictorial.html
 *  Building R Packages with Devtools (Jiddu Alexander) http://www.jiddualexander.com/blog/r-package-building/
 *  Developing R packages (Jeff Leek) https://github.com/jtleek/rpackages
 *  R Package Tutorial (Colautti Lab) https://colauttilab.github.io/RCrashCourse/Package_tutorial.html
 *  Instructions for creating your own R package (MIT) http://web.mit.edu/insong/www/pdf/rpackage_instructions.pdf
 *  How to Create and Distribute an R Package (Shian Su) https://medium.com/@shiansu/how-to-create-and-distribute-an-r-package-a296217435dc
 *  R Forwards Package Workshop (Chicago, February 23, 2019) https://github.com/forwards/workshops/tree/master/Chicago2019
 *  Write your own R package (UBC STAT 545) http://stat545.com/packages00_index.html
 *  R package workshop https://combine-australia.github.io/r-pkg-dev/
 
  





## 'Stable-enough' packages within `{unhcrverse}`

The list of mature projects includes set of packages to ease the production of statistical evidence and data stories. Even if not installable throught the main CRAN repository, those **Packages** are expected to follow the standards outlined in the [CRAN policies](https://cran.r-project.org/web/packages/policies.html).

> However as those packages heavily relies on the benevolent time of the maintainers and, while wiating for for systematic third-party code review process set up, each those package might not be (yet!) fully CRAN-Check compliant... 



`{unhcrverse}` includes packages related to 


 
 1. __Design and branding__ such as [unhcrdown](https://github.com/vidonne/unhcrdown): UNHCR templates for R Markdown,   [unhcrthemes](https://github.com/vidonne/unhcrthemes): UNHCR branded theme for ggplot2 and data visualization colour palettes, 
 
 
 2. __Data Access__ such as [unhcrdatapackage](https://github.com/edouard-legoupil/unhcrdatapackage): Use UNHCR Open data, [ridl](https://github.com/dickoa/ridl): API to connect to internal data source, [robotoolbox](https://github.com/dickoa/obotoolbox): API to connect to Kobo Data Collection Server, [rhdx](https://github.com/dickoa/hcrdata): API to connect to Humanitarian Data Exchange and [popdata](https://github.com/PopulationStatistics/popdata): API to connect to the system for official statistic curation


 3. __Survey Data Processing__ such as [HighFrequencyChecks](https://github.com/edouard-legoupil/HighFrequencyChecks/): Perform High Frequency Check or  [kobocruncher](https://github.com/edouard-legoupil/kobocruncher): Process data crunching for survey dataset or [XlsFormPrettyPrint](https://github.com/edouard-legoupil/XlsFormPrettyPrint): to work with multiple xlsform

You can install them all with the following:

```{r unhcrverse, eval = FALSE, include=TRUE}
install.packages("pak")
pak::pkg_install("unhcRverse/unhcRverse")
unhcrverse::unhcrverse()
```




## Deprecated projects

Any stable project where there is no maintainer working with UNHCR, or when the scope got filled by other better packages will be placed in the list of deprecated packages.

For instance: 

 *   [koboloader](https://unhcr.github.io/koboloadeR/docs/) got deprecated by  [kobocruncher](https://github.com/edouard-legoupil/kobocruncher)
 
 *  [unhcrstyle](https://unhcr-web.github.io/unhcRstyle/docs/) got deprecated by [unhcrdown](https://github.com/vidonne/unhcrdown) and  [unhcrthemes](https://github.com/vidonne/unhcrthemes)



```{r echo=FALSE}


# <!--  
# 
#   rec1 [label = 'Data \n Extraction & \n Shaping']
# * [popdata](https://github.com/unhcRverse/opdata): API to connect to the system for official statistic curation
# * [ridl](https://github.com/unhcRverse/ridl): API to connect to internal data source
# * [robotoolbox](https://github.com/unhcRverse/hcrdata): API to connect to Kobo Data Collection Server
# * [rhdx](https://github.com/unhcRverse/hcrdata): API to connect to Humanitarian Data Exchange
# 
#   rec2 [label = 'Data \n Quality \n Check']
# * [HighFrequencyChecks](https://github.com/unhcRverse/HighFrequencyChecks/): Perform High Frequency Check
# 
#   rec3 [label =  'Data Exploration /\n  Chart Library']
# * [unhcrdatapackage](https://github.com/unhcRverse/unhcrdatapackage): Use UNHCR Open data
# * [koboloadeR](https://github.com/unhcRverse/kobocruncher/): Process data crunching for survey dataset
# 
#   rec4 [label = 'Branding: \n Report Template &\n  Chart Style']
# * [unhcrdown](https://github.com/unhcRverse/unhcrdown): UNHCR templates for R Markdown
# * [unhcrthemes](https://github.com/unhcRverse/unhcrthemes): UNHCR branded theme for ggplot2 and data visualization colour palettes
# -->

#  graph [layout = dot, rankdir = TB]
# label= "Comic Sans MS"

  # layout = neato

# 'transparent'
unhcrverse <- DiagrammeR::grViz("digraph {
  bgcolor = transparent
  compound=true;
  graph [layout = dot, 
         rankdir = TB,
         fontsize= 46, 
        color = transparent,
        fontname='Lato',
        fontcolor='#FAEB00']
  node [shape = hexagon, 
        fixedsize = true,
        width = 2.5,
        height = 2.5,
        color = '#FAEB00',
        penwidth = 4,
        fontsize=17, 
        fontname='Lato',
        fontcolor='#ffffff']  
        
  edge [color = '#fffff',
        penwidth = 3]  
        
  subgraph cluster_unhcrverse {
         label = 'unhcRverse Stable-enough packages: Categories & Dependencies'
    
  
      rec1 [label = 'Data \n Extraction & \n Shaping']
      subgraph cluster_1 {
            label = ' '  
            rec1 -> rec5
            rec1 -> rec6
            rec1 -> rec7
            rec1 -> rec8   ;
            color='#ffffff'; }
    
      rec2 [label = 'Data \n Quality \n Check']
      subgraph cluster_2 {
            label = ' '
            rec2 -> rec9 
            rec2 -> rec99  ;
            color='#ffffff';  }
      
      rec3 [label =  'Data Exploration /\n  Chart Library']
      subgraph cluster_3 {
            label = ' ';
            rec3 -> rec10;
            rec3 -> rec11;
            rec3 -> rec112 ;
            rec3 -> rec113 ;
            color='#ffffff'; }
      
      rec4 [label = 'Branding: \n Report Template &\n  Chart Style']
      subgraph cluster_4 {
            label = ' ';
            rec4 -> rec12;
            rec4 -> rec13  ;
            color='#ffffff'; }

  }   
  
  
  rec5 [label = 'popdata:\n pull from \n popdata.unhcr.org']
  rec6 [label = 'ridl:\n pull/push from \n ridl.unhcr.org']
  rec7 [label = 'robotoolbox:\n pull from \n kobo.unhcr.org']
  rec8 [label = 'rhdx:\n pull from \n data.humdata.org']
  
  
  rec9 [label = 'High\n FrequencyChecks:\n Data collection quality']
  rec99 [label = 'Statistical\n Quality\n Assurance:\n SQAF']
  
  
  rec10 [label = 'unhcrdatapackage:\n Population Statistics']
  rec11 [label = 'iati:\n UNHCR activities &\n Budget']
  rec112 [label = 'kobocruncher:\n Automatic \n data exploration \n for surveys']
  rec113 [label = 'progresAnalysis:\n Registration analysis']
  
  rec12 [label = 'unhcrdown:\n Report template']
  rec13 [label = 'unhcrthemes:\n Ggplot theme']
  
  
   rec5  -> rec99
   
   rec7  -> rec9
   rec9  -> rec6
   
   rec6  -> rec113
   
   
   rec12  -> rec112
   rec13  -> rec112
   
   
   rec12  -> rec113
   rec13  -> rec113
   
   rec12  -> rec11
   rec13  -> rec11
   
   rec12  -> rec10
   rec13  -> rec10
  
  }",
  height = 550)

#3require(magrittr)
#require(DiagrammeR)
#require(DiagrammeRsvg)
#require(xml2)

# unhcrverse |>
#   DiagrammeRsvg::export_svg() |>
#   xml2::read_xml() |>
#   xml2::write_xml("unhcrverse.svg")
# 
# 
# unhcrverse |>
#   DiagrammeRsvg::export_svg() |>
#   charToRaw() |>
#   rsvg::rsvg_png("unhcrverse.png")

```