---
title: "Package Ecosystem"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### 'Stable-enough' packages within `unhcRverse`

The list of mature projects includes set of packages to ease the production of statistical evidence and data stories. Even if not installable throught the main CRAN repository, those **Packages** are expected to follow the standards outlined in the [CRAN policies](https://cran.r-project.org/web/packages/policies.html). However as those packages heavily relies on the benevolent time of the maintainer and in the absence of systematic third-party code review, each those package might not be fully CRAN-Check compliant... 

* [unhcrdown](https://github.com/unhcRverse/unhcrdown): UNHCR templates for R Markdown
* [unhcrthemes](https://github.com/unhcRverse/unhcrthemes): UNHCR branded theme for ggplot2 and data visualization colour palettes
* [unhcrdatapackage](https://github.com/unhcRverse/unhcrdatapackage): Use UNHCR Open data
* [ridl](https://github.com/unhcRverse/ridl): API to connect to internal data source
* [robotoolbox](https://github.com/unhcRverse/hcrdata): API to connect to Kobo Data Collection Server
* [rhdx](https://github.com/unhcRverse/hcrdata): API to connect to Humanitarian Data Exchange
* [popdata](https://github.com/unhcRverse/opdata): API to connect to the system for official statistic curation
* [HighFrequencyChecks](https://github.com/unhcRverse/HighFrequencyChecks/): Perform High Frequency Check
* [koboloadeR](https://github.com/unhcRverse/koboloadeR/): Process data crunching for survey dataset

You can install them all with the following:

```{r unhcrverse, eval = FALSE, include=TRUE}
## unhcrdown
remotes::install_github("unhcRverse/unhcrdown")
## unhcrthemes
remotes::install_github("unhcRverse/unhcrthemes")
## unhcrdatapackage
remotes::install_github('unhcRverse/unhcrdatapackage')
## popdata
remotes::install_github('unhcRverse/popdata')
## ridl
remotes::install_github('unhcRverse/ridl')
## robotoolbox
remotes::install_github('unhcRverse/robotoolbox')
## rhdx
remotes::install_github('unhcRverse/rhdx')
## HighFrequencyChecks
remotes::install_github('unhcRverse/HighFrequencyChecks')
## koboloadeR
remotes::install_github('unhcRverse/koboloadeR')
```




### Deprecated projects

Any stable project where there is no maintainer working with UNHCR will be placed as deprecated.



```{r}


# <!--  
# 
#   rec1 [label = 'Data \n Extraction & \n Shaping']
# * [popdata](https://github.com/unhcRverse/opdata): API to connect to the system for official statistic curation
# * [ridl](https://github.com/unhcRverse/ridl): API to connect to internal data source
# * [robotoolbox](https://github.com/unhcRverse/hcrdata): API to connect to Kobo Data Collection Server
# * [rhdx](https://github.com/unhcRverse/hcrdata): API to connect to Humanitarian Data Exchange
# 
#   rec2 [label = 'Data \n Quality \n Check']
# * [HighFrequencyChecks](https://github.com/unhcRverse/HighFrequencyChecks/): Perform High Frequency Check
# 
#   rec3 [label =  'Data Exploration /\n  Chart Library']
# * [unhcrdatapackage](https://github.com/unhcRverse/unhcrdatapackage): Use UNHCR Open data
# * [koboloadeR](https://github.com/unhcRverse/kobocruncher/): Process data crunching for survey dataset
# 
#   rec4 [label = 'Branding: \n Report Template &\n  Chart Style']
# * [unhcrdown](https://github.com/unhcRverse/unhcrdown): UNHCR templates for R Markdown
# * [unhcrthemes](https://github.com/unhcRverse/unhcrthemes): UNHCR branded theme for ggplot2 and data visualization colour palettes
# -->

#  graph [layout = dot, rankdir = TB]
# label= "Comic Sans MS"

  # layout = neato

# 'transparent'
unhcrverse <- DiagrammeR::grViz("digraph {
  bgcolor = transparent
  compound=true;
  graph [layout = dot, 
         rankdir = TB,
         fontsize= 46, 
        color = transparent,
        fontname='Lato',
        fontcolor='#FAEB00']
  node [shape = hexagon, 
        fixedsize = true,
        width = 2.5,
        height = 2.5,
        color = '#FAEB00',
        penwidth = 4,
        fontsize=17, 
        fontname='Lato',
        fontcolor='#ffffff']  
        
  edge [color = '#fffff',
        penwidth = 3]  
        
  subgraph cluster_unhcrverse {
         label = 'unhcRverse Stable-enough packages: Categories & Dependencies'
    
  
      rec1 [label = 'Data \n Extraction & \n Shaping']
      subgraph cluster_1 {
            label = ' '  
            rec1 -> rec5
            rec1 -> rec6
            rec1 -> rec7
            rec1 -> rec8   ;
            color='#ffffff'; }
    
      rec2 [label = 'Data \n Quality \n Check']
      subgraph cluster_2 {
            label = ' '
            rec2 -> rec9 
            rec2 -> rec99  ;
            color='#ffffff';  }
      
      rec3 [label =  'Data Exploration /\n  Chart Library']
      subgraph cluster_3 {
            label = ' ';
            rec3 -> rec10;
            rec3 -> rec11;
            rec3 -> rec112 ;
            rec3 -> rec113 ;
            color='#ffffff'; }
      
      rec4 [label = 'Branding: \n Report Template &\n  Chart Style']
      subgraph cluster_4 {
            label = ' ';
            rec4 -> rec12;
            rec4 -> rec13  ;
            color='#ffffff'; }

  }   
  
  
  rec5 [label = 'popdata:\n pull from \n popdata.unhcr.org']
  rec6 [label = 'ridl:\n pull/push from \n ridl.unhcr.org']
  rec7 [label = 'robotoolbox:\n pull from \n kobo.unhcr.org']
  rec8 [label = 'rhdx:\n pull from \n data.humdata.org']
  
  
  rec9 [label = 'High\n FrequencyChecks:\n Data collection quality']
  rec99 [label = 'Statistical\n Quality\n Assurance:\n SQAF']
  
  
  rec10 [label = 'unhcrdatapackage:\n Population Statistics']
  rec11 [label = 'iati:\n UNHCR activities &\n Budget']
  rec112 [label = 'kobocruncher:\n Automatic \n data exploration \n for surveys']
  rec113 [label = 'progresAnalysis:\n Registration analysis']
  
  rec12 [label = 'unhcrdown:\n Report template']
  rec13 [label = 'unhcrthemes:\n Ggplot theme']
  
  
   rec5  -> rec99
   
   rec7  -> rec9
   rec9  -> rec6
   
   rec6  -> rec113
   
   
   rec12  -> rec112
   rec13  -> rec112
   
   
   rec12  -> rec113
   rec13  -> rec113
   
   rec12  -> rec11
   rec13  -> rec11
   
   rec12  -> rec10
   rec13  -> rec10
  
  }",
  height = 550)

#3require(magrittr)
#require(DiagrammeR)
#require(DiagrammeRsvg)
#require(xml2)

unhcrverse |>
  DiagrammeRsvg::export_svg() |>
  xml2::read_xml() |>
  xml2::write_xml("unhcrverse.svg")


unhcrverse |>
  DiagrammeRsvg::export_svg() |>
  charToRaw() |>
  rsvg::rsvg_png("unhcrverse.png")

```