---
title: "Creating functions, a first step into R programming"
subtitle: "Illustration with {unhcrplot}"
date: <a href="index.html"><i class="fa fa-indent  fa-fw fa-2x"></i></a>
output: 
  unhcrdown::html_slides:
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
---
 

# Learning objectives

  * Understand Why Functions are needed
  
  * Learn to transition from Scripts to Functions
  
  * Call Custom Function
  
  * Function Documentation
  
  * Plotting Functions
  
  * Function Parameter

 
---

## Why Functions?  

Code Reusability

  * Functions allow you to reuse code.
  * Write a function once and use it multiple times.

Modularity

  * Functions help in breaking down complex tasks into smaller, manageable pieces.
  * Easier to debug and maintain.

Collaboration

  * Functions make it easier to collaborate with others.
  * Share functions instead of long scripts.

---

## Transition from Scripts to Functions


.pull-left[

Script Example

```{r}
# Script to calculate the mean of a numeric vector
data <- c(1, 2, 3, 4, 5)
mean_value <- mean(data)
print(mean_value)
```


]

.pull-right[

 

Function Example

```{r}

# Function to calculate the mean of a numeric vector
calculate_mean <- function(data) {
  mean_value <- mean(data)
  return(mean_value)
}

data <- c(1, 2, 3, 4, 5)
mean_value <- calculate_mean(data)
print(mean_value)
```

]

---

## Key Differences


.pull-left[

  * Functions encapsulate code into reusable units.
  * Parameters allow customization of function behavior.


]

.pull-right[

Function Structure

```{r}

function_name <- function(parameter1, parameter2, ...) {
  # Function body
  # Perform operations
  return(result)
}
```
]

---

## Example: Custom Function


.pull-left[

 1. Takes an input

 2. Apply a transformation

 3. Generates an output


]

.pull-right[


```{r}

# Custom function to calculate the median of a numeric vector
calculate_median <- function(data) {
  median_value <- median(data)
  return(median_value)
}
```
]

---

## Calling Custom Function

```{r}

data <- c(1, 2, 3, 4, 5)
median_value <- calculate_median(data)
print(median_value)
```


---

## Import Data 


from Excel - [`readxl`](https://readxl.tidyverse.org/),


from Text: [`readr`](https://readr.tidyverse.org/),  etc.


---

## Why Document Functions?

  * Documentation provides information on how to use a function.
  
  * Helps other users (including your future self) understand the purpose and usage of the function.
  
  * Standard practice for sharing code.

R Documentation

  * Use roxygen2 or devtools for generating R documentation.
  * Add details about parameters, return values, and examples.
  
Code Style and Consistency 

  * Follow modern and consistent code formatting.
  * Use tools like styler for code formatting.

---

## Example Documentation

```{r}

#' Calculate the median of a numeric vector.
#'
#' This function takes a numeric vector and returns its median value.
#'
#' @param data A numeric vector.
#'
#' @return The median of the input vector.
#'
#' @examples
#' data <- c(1, 2, 3, 4, 5)
#' median_value <- calculate_median(data)
#'
#' @export
calculate_median <- function(data) {
  median_value <- median(data)
  return(median_value)
}
```


---

## Use Modular Code
 

  * Modular code is organized into small, reusable components.
  * Makes your code more readable and maintainable.
  * Reduces redundancy and duplication.

Recurrent Plots

  * Recurrent plots are plots used multiple times in your code or projects.

---

## What is a Plotting Function?

  * A plotting function is a custom R function that generates a specific plot.
  * Encapsulates the plot's code and logic.

Advantages

  * Reusability
  * Consistency
  * Code organization

```{r}

# Define a plotting function
# plot_histogram <- function(data, variable) {
#   ggplot(data, aes(x = {{ variable }})) +
#     geom_histogram(binwidth = 1, fill = "lightblue") +
#     labs(title = paste("Histogram of", deparse(substitute(variable))))
# }
```


---

## Function Parameters

  * Plotting functions can accept parameters to customize the plot.
  * Parameters allow you to create dynamic and flexible plots.

Customization Example

```{r}

# Customize the plotting function
#plot_histogram(mtcars, mpg)
```
 
Plot Customization

  * Allow users to customize plots by providing parameters.
  * Parameters can control titles, colors, labels, etc.

Example with Custom Parameters

```{r}

# Customize plot appearance
# plot_histogram(mtcars, mpg, title = "Custom Histogram", fill_color = "lightgreen")
```


---

## Organizing Plotting Functions

  * Create a directory or package to store your plotting functions.
  * Name files descriptively and include documentation.

---

## Advanced Features

  * Saving plots to svg or png
  * Creating dynamic dashboards
  * Integrate your plotting functions into Shiny apps for interactive visualization.

---

## Key Takeaways

  * Functions are essential for code re-usability and modularity.  
  
  * Transition from scripts to functions for better code organization. 
  
  * Documenting functions is crucial for effective collaboration and code sharing.  
  
  * Modular code with plotting functions improves code organization and re-usability.  
 
  * Function parameters allow customization of plots.  
 
  * Building a directory of plotting functions streamlines your work.

---
class: inverse, center, middle

# Thank you

### Questions?


[post Feedback here](https://github.com/unhcRverse/unhcrverse/issues/new?assignees=&labels=enhancement&projects=&template=comment_prex_2_tidyverse.md&title=%5Blearn%5D)

<a href="index.html"><i class="fa fa-indent  fa-fw fa-2x"></i></a>


---

# Reference

 - [Build Advanced Charts in R](https://wd3.myworkday.com/unhcr/learning/course/efee8410b2b410017044140f1aeb0001?type=9882927d138b100019b928e75843018d)
 
 - [Data Wrangling in R](https://wd3.myworkday.com/unhcr/learning/course/046437bef6c810195cf76ead31590002?type=9882927d138b100019b928e75843018d)

 - [Code Clinic: R](https://wd3.myworkday.com/unhcr/learning/course/046437bef6c810195cfbd7b02bd90006?type=9882927d138b100019b928e75843018d)

 - [Coding Exercises: R Data Science](https://wd3.myworkday.com/unhcr/learning/course/046437bef6c810195cf7768256ed0000?type=9882927d138b100019b928e75843018d)

 - [R for Data Science](https://r4ds.had.co.nz/), 

 - [Advanced R Programming](https://adv-r.hadley.nz/)
 
---

# Excercise

Use the char your previously created and turn it into a function with parameters
